Current and future constraints on primordial magnetic fields

### 2017-07-24 response_rnd2.tex

The reviewer's comments are shown in \reviewer{blue,} our response is in black, and the revised text in the paper is in \diff{green}. \\

\reviewer{1. The abstract does not mention that the constraints are derived specifically on the scale-invariant PMF, nor it mentions what $A_{PMF}$ is and how it is related to the PMF strength.}

We have restated the abstract to be more clear:
\diff{We reduce twofold the 95\% CL upper limit on the CMB anisotropy power due to a nearly-scale-invariant PMF, with an allowed B-mode power at $\ell=1500$ of $D_{\ell=1500}^{BB} < 0.071 \mu K^2$ for \planck{} versus $D_{\ell=1500}^{BB} < 0.034 \mu K^2$ for the combined dataset. }
\\\\
We have also added supporting text to the Results section:
\diff{To facilitate a physical interpretation, we can restate these limits as an allowed B-mode power  at $\ell=1500$, which is near the peak of the PMF contribution. 
The equivalent limits on the allowed B-mode power are $D_{\ell=1500}^{BB} < 0.071 \mu K^2$ for \planck{}-alone and $D_{\ell=1500}^{BB} < 0.034 \mu K^2$ when the ground-based experiments are added. }
\\\\
\reviewer{2. The drafts says "... scalar contributions are not implemented in the code", referring to the Zucca et el modification of CAMB. That is not true. Instead, it should say that the authors chose not to use the scalar contributions.}

We have struck ``in the code" from the sentence.
\\\\
\reviewer{3. To a casual reader it would seem mysterious that the tensor mode amplitude depends on beta as a power of 1.9 that had to be found by fitting. The tensor mode spectrum is simply multiplied by a factor of $[const+ln(a_nu/a_PMF)]^2$. It would make more sense to simply define beta with the constant included. At least, the authors should explain that its 1.9 and not 2 because of the way beta was defined.}
\\\\
We prefer to keep the current definition of $\beta$ rather than introducing an additional additive constant, as it parallels the definitions used in previous constraints. For instance, Zucca et al 2016 define $\beta$ around Table II as $\beta = log_{10}(\tau_\nu/\tau_B)$; Planck 2015 results.

We have edited the language to better reflect the expectation:
\diff{The power in the tensor modes also depends on the timing of PMF generation relative to neutrino decoupling, and is expected to scale approximately as $\beta^2$ \citep{lewis04,shaw10b}.   
We double-check this expectation for $\beta$ in the range 2 to 20 using the  \citet{zucca16} code, and find the computed tensor power scales as $\beta^{1.9}$. }

### 2017-08-09 pmf.tex
fixed typos

We use MCMC methods to determine the parameter constraints reported in this work. 
The results are calculated using  the {\textsc CosmoMC}\footnote{http://cosmologist.info/cosmomc (July 2015 version)} package \citep{lewis02b}. 
CosmoMC invokes  CAMB\footnote{http://camb.info}  \citep{lewis00} to calculate the CMB power spectrum for each set of cosmological parameters. 
Although CosmoMC and CAMB have a partial implementation of PMFs, we choose to adopt a simpler, fast, template-based calculation for the PMF-sourced power. 
We have adapted CosmoMC to add a scaled version of a PMF template to all four CMB power spectra: TT, TE, EE, and BB, where the scale factor is \apmf. 
\be
\changed{C_\ell = C_\ell^{\rm{CAMB}} + \apmf \left[\cpmfvec + \left(\frac{\beta}{20}\right)^{1.9}  \cpmftens \right]}\label{clmodel}
\ee
\changed{The calculation of the PMF templates, \cpmfvec{} and \cpmfvec{}, and the motivation for the $\beta$ scaling are described in \S\ref{sec:template}. 
The parameter $\beta = {\rm ln}(a_{\nu}/a_{\rm{PMF}})$ relates the timing of neutrino decoupling and the generation of the PMF, where $a_x$ represents the scale factor at the respective events. }
Note that no other effect of  PMFs is considered; this is one reason we only use CMB data.

@@

Given that we are still in the era of upper limits, we \changed{fix the value of $n_B$ instead of exploring the full 3-dimensional PMF parameter space. 
For the real data, we focus on nearly-scale-invariant ($n_B = -2.9$) PMF template in this work due to its connection to inflation.  
Quantum mechanical zero-point fluctuations during inflation can generate an approximately scale-invariant PMF with a spectral index close to $n_B = -3$ \citep[e.g.,][]{turner88, mack01,2008PhRvD..78f3012K}. %\footnote{We select $n_B = -2.9$ as opposed to another number near -3 to facilitate comparison to other PMF papers, while avoiding the singularity at -3.}
In contrast, causal PMFs with bluer spectra require later phase transitions or other new physics after inflation \citep[e.g.,][]{durrer03, subramanian16}. 
We also consider $n_B\,=\,-1$ and $n_B\,=\,2$ for the forecasts in \S\ref{sec:forecasts}. 
For each value of $n_B$, we use publicly released modifications to CAMB from \citet{zucca16} to calculate a PMF template for the CMB TT, TE, EE and BB power spectra at fixed parameters: \bpmf = 2.5\,nG and $\beta=20.72$ ($a_{\nu}/a_{\rm{PMF}} = 10^9$).} 
The calculated templates include tensor and vector PMF contributions; scalar contributions are not implemented. 
The lack of scalar terms should not matter for CMB polarization data. 
Temperature data is not the main focus of this work, and  and we find our \planck{} results  to be consistent with \citet{planck15-19}, suggesting the impact of the scalar terms is minor. 
The PMF templates are plotted against other signals in the CMB B-mode power spectrum in Fig.~\ref{fig:pmf-bb} \changed{and against each other in Fig.~\ref{fig:pmf-nb}. }

>>>>

We use MCMC methods to determine the parameter constraints reported in this work. 
The results are calculated using  the {\textsc CosmoMC}\footnote{http://cosmologist.info/cosmomc (July 2015 version)} package \citep{lewis02b}. 
CosmoMC invokes  CAMB\footnote{http://camb.info}  \citep{lewis00} to calculate the CMB power spectrum for each set of cosmological parameters. 
Although CosmoMC and CAMB have a partial implementation of PMFs, we choose to adopt a simpler, fast, template-based calculation for the PMF-sourced power. 
We have adapted CosmoMC to add a scaled version of a PMF template to all four CMB power spectra: TT, TE, EE, and BB, where the scale factor is \apmf. 
\be
\changed{C_\ell = C_\ell^{\rm{CAMB}} + \apmf \left[\cpmfvec + \left(\frac{\beta}{20}\right)^{1.9}  \cpmftens \right]}\label{clmodel}
\ee
\changed{The calculation of the PMF templates, \cpmfvec{} and \cpmftens{}, and the motivation for the $\beta$ scaling are described in \S\ref{sec:template}. 
The parameter $\beta = {\rm ln}(a_{\nu}/a_{\rm{PMF}})$ relates the timing of neutrino decoupling and the generation of the PMF, where $a_x$ represents the scale factor at the respective events. }
Note that no other effect of  PMFs is considered; this is one reason we only use CMB data.

@@

Given that we are still in the era of upper limits, we \changed{fix the value of $n_B$ instead of exploring the full 3-dimensional PMF parameter space. 
For the real data, we focus on nearly-scale-invariant ($n_B = -2.9$) PMF template in this work due to its connection to inflation.  
Quantum mechanical zero-point fluctuations during inflation can generate an approximately scale-invariant PMF with a spectral index close to $n_B = -3$ \citep[e.g.,][]{turner88, mack01,2008PhRvD..78f3012K}. %\footnote{We select $n_B = -2.9$ as opposed to another number near -3 to facilitate comparison to other PMF papers, while avoiding the singularity at -3.}
In contrast, causal PMFs with bluer spectra require later phase transitions or other new physics after inflation \citep[e.g.,][]{durrer03, subramanian16}. 
We also consider $n_B\,=\,-1$ and $n_B\,=\,2$ for the forecasts in \S\ref{sec:forecasts}. 
For each value of $n_B$, we use publicly released modifications to CAMB from \citet{zucca16} to calculate a PMF template for the CMB TT, TE, EE and BB power spectra at fixed parameters: \bpmf = 2.5\,nG and $\beta=20.72$ ($a_{\nu}/a_{\rm{PMF}} = 10^9$).} 
The calculated templates include tensor and vector PMF contributions; scalar contributions are not implemented. 
The lack of scalar terms should not matter for CMB polarization data. 
Temperature data is not the main focus of this work, and we find our \planck{} results  to be consistent with \citet{planck15-19}, suggesting the impact of the scalar terms is minor. 
The PMF templates are plotted against other signals in the CMB B-mode power spectrum in Fig.~\ref{fig:pmf-bb} \changed{and against each other in Fig.~\ref{fig:pmf-nb}. }
